Collaborative Test-Driven Scientific Model Validation

### 2013-11-23 sciunit-icse14/sciunit-icse14.tex
Started describing unit testing
more progress on intro

Quantitative models are described and validated using a social process: the scientific publication system. For a model to be accepted by the scientific community, scientists must write a paper describing it and providing  evidence that it predicts some quantity of interest more accurately than previous models, or that the tradeoff it makes between accuracy and complexity may be useful to the community \cite{box}. During the \emph{peer review} process, other members of the relevant community are tasked with ensuring that the validation criteria used (e.g. a $p$-value) are strong enough and that all relevant data and competing models were fully considered, drawing on knowledge of statistical methods and the prior literature. Publishing a paper is one of the primary motivators for most scientists \cite{howison2011scientific}.

Quantitative scientific modeling and software development have much in common. Indeed, quantitative models are increasingly implemented as computer programs. In some cases, the program \emph{is} the model (e.g. complex simulations). The peer review process is similar in many ways to the \emph{code review} process used in many development teams. During code review, team members look for errors, enforce style and architectural guidelines and check that the code is \emph{valid} (i.e. that achieves its intended goal \cite{validity}) before it can be committed to the primary source code repository \cite{codereview}.

Code review is reasonably effective \cite{creffectiveness1,creffectiveness2}. However, to be most effective reviewers must expend considerable effort \cite{effort}, and code review is most effective for resolving  issues related to software architecture \cite{evolvability}. 
For these reasons, most software development teams supplement code reviews with a number of more {automated} approaches to verification and validation. One of the most widely-adopted approaches is \emph{unit testing}. In brief, unit tests are functions that check a single component of a program against a single well-defined correctness criterion \cite{beck2003}. A suite of such tests can be seen as a partial specification of a program or component.  Test suites complement code review by allowing developers to answer questions like these more easily:
\begin{enumerate}
\item How much development progress has been made? Which functionality has not yet been adequately implemented? How valuable is a code contribution?
\item What modes of usage does the team consider most important? How does the development team measure success?
\item Does a code contribution cause \emph{regressions} in other parts of a program?
\end{enumerate}
Scientists ask analagous questions:
\begin{enumerate}
\item When new experimental observations are made, scientists want to determine the impact they have on the validity of previously-developed models.
\end{enumerate}
Answering these questions is often difficult using scientific publications alone. Each publication is frozen in time, so its examination of validity cannot consider the most recent body of data, or modern validation criteria. Discovering these validation criteria, as well as precisely characterizing open modeling problems  requires an encyclopedia knowledge of the literature. In many areas, the number of publications generated every year can overwhelm even the most conscientious and experienced scientists \cite{jinha_article_2010}. A common complaint of statisticians is that scientists are not aware of best practices and that community standards evolve too slowly \cite{grumpystatisticians}.

Unfortunately, there are few alternatives to comprehensive literature review available when scientists need to answer questions like these today.%Anecdotally, one of the greatest fears of new scientists is that a problem that they are working on has already been solved and published.

@@

-------

\begin{itemize}
\item Which models are capable of predicting the quantities I am interested in?
\item Which metrics should be used to evaluate the goodness-of-fit between these models and  data?
\item How well do these models perform, as judged to these metrics, given currently available data?
\item What other quantities can and can't these models predict?
\item What observations are not adequately explained by any available model?
\end{itemize}

>>>>

Quantitative models are validated by \emph{peer review}. For a model to be accepted by the scientific community, scientists must submit a paper describing it and providing  evidence that it predicts some quantity of interest more accurately than previous models, or that it makes a desirable tradeoff  between accuracy and complexity \cite{box}. Other members of the relevant community are then tasked with ensuring that validity was measured properly and that relevant data and competing models were adequately considered, drawing on knowledge of statistical methods and the prior literature. Publishing is a primary motivator for most scientists \cite{howison2011scientific}.

Quantitative scientific modeling and software development have much in common. Indeed, quantitative models are increasingly  implemented as computer programs and in some cases, the program \emph{is} the model (e.g. complex simulations). The peer review process for papers is similar in many ways to the \emph{code review} process used in many development teams, where team members look for errors, enforce style and architectural guidelines and check that the code is \emph{valid} (i.e. that achieves its intended goal \cite{validity}) before permitting it to be committed to the primary source code repository \cite{codereview}.

Code review can be quite effective \cite{creffectiveness1,creffectiveness2}, but this requires that developers  expend considerable effort \cite{effort}. Code review is also most effective for resolving  issues related to software architecture \cite{evolvability}. 
Most development teams thus supplement code reviews with {automated} approaches to verification and validation, the most widely-used  of which is \emph{unit testing} \cite{beck2003}. In brief, unit tests are functions that check a single portion of a program against a single well-defined correctness criterion. A suite of such tests can be seen as a partial specification of a program or component.  Test suites complement code review by allowing developers to answer questions like these more easily:
\begin{enumerate}
\item Which functionality has been adequately implemented? What remains to be done?
%\item How much progress does an individual code contribution make? 
\item What modes of usage does the team consider most important? How does the team measure correctness?
\item Does a code contribution cause \emph{regressions} in other parts of a program?
\end{enumerate}
Scientists ask analagous questions:
\begin{enumerate}
\item Which observations are already well-explained by existing models? What is the state-of-the-art? What are the open modeling problems of interest?
%\item How much better is a new model than the state of the art? What are the open problems in my field?
\item What are the contemporary community standards for measuring goodness-of-fit?
\item How do newly-made experimental observations impact the validity of previous models?
\end{enumerate}
But while software engineers can rely on a program's test suite, scientists today must extract this information from a body of scientific publications. This is increasingly difficult. Each publication focuses on just one model and is frozen in time, so it does not consider the latest experimental data or statistical methods. Discovering, precisely characterizing and comparing several models to discover the state of the art, and finding open modeling problems, often requires an encyclopedic knowledge of the literature. Senior scientists can attempt to summarize this with review papers, but in many areas, the number of publications generated every year can be overwhelming \cite{jinha_article_2010}, and comprehensive reviews are published relatively infrequently for any given area. Statisticians commonly complain that scientists are not following best practices and that community standards evolve too slowly because canonical papers or popular reviews advocate outdated methods \cite{grumpystatisticians}. One might compare this to a  software development team relying primarily on carefully-prepared minutes of code review sessions, peppered with cross-references to previous minutes, to understand and validate a complex software project.

We suggest that scientific model validation could benefit from the adoption of a test-driven methodology. However, scientific communities differ from typical development teams in several important ways.

- Unit tests are typically pass/fail, while goodness-of-fit is typically measured more granularly. 
- Different communities use different goodness-of-fit measures. These continuously evolve as, for example, statisticians develop better methods.
- Professional software developers are used to testing, while scientists are professional end-users and rarely use testing, even for verification.

%Anecdotally, one of the greatest fears of new scientists is that a problem that they are working on has already been solved and published.

@@
